\large \textbf{Title:} The role of learning in the evolution of status signalling: a modeling approach

### 2023-09-19 manuscript/manuscript_1.0.tex
Elimate files no longer needed for the publication

\begin{document}
\maketitle
\begin{flushleft}
\end{flushleft}
\begin{abstract}
Adaptive behavioural responses often depend on qualities of the
interacting partner of an individual. For example, when competing for
resources, an individual might be better off escalating fights with
individuals of lower quality, while restraining from fighting
individuals of higher quality. Communication systems involving signals
of quality allow individuals to reduce uncertainty regarding the
fighting ability of their partners and make more adaptive behavioral
decisions. However, dishonest individuals can destabilize such
communications systems. An open question is whether cognitive
mechanisms, such as learning, can maintain the honesty of signals, thus
favoring their evolutionary stability. We present evolutionary
simulations where individuals can produce a signal proportional to their
quality and learn along their lifetime the best response to the signals
emitted by their peers. Our simulations replicate previous results where
the handicap principle mediates the evolution of signals. In our
simulations learning on the receiver side can mediate the evolution of
signals of quality on the sender side. When the cost of the signal is
proportional to the quality of the sender, all individuals in
populations are honest signalers. In contrast to traditional models
which predict the absence of signals when the cost is not proportional
to the quality of the signaler, our model revealed that learning
facilitates the evolution of a polymorphism in which populations
comprise both honest and dishonest signalers. We argue that learning may
have a role in the evolution and dynamics of a wide range of
communication systems and more generally in behavioral responses.
\end{abstract}
\subsection*{Lay summary}
Biological signals allow individuals to respond appropietly. For
example, a singal of fighting hability would allow peers to decide
whether to fight another individual. But, what prevents a weaklings from
faking to be strong? Using a theoretical approach we assess whether
learning has a role in maintaining signals honest. We find that when
individuals learn how best to respond to a signal, populations are
composed of a mixed of honest and dishonest signals.

>>>>

\begin{document}
\maketitle
\begin{flushleft}
\end{flushleft}
\begin{abstract}
Adaptive behavioural responses often depend on qualities of the
interacting partner of an individual. For example, when competing for
resources, an individual might be better off escalating fights with
individuals of lower quality, while restraining from fighting
individuals of higher quality. Communication systems involving signals
of quality allow individuals to reduce uncertainty regarding the
fighting ability of their partners and make more adaptive behavioral
decisions. However, dishonest individuals can destabilize such
communications systems. An open question is whether cognitive
mechanisms, such as learning, can maintain the honesty of signals, thus
favoring their evolutionary stability. We present evolutionary
simulations where individuals can produce a signal proportional to their
quality and learn along their lifetime the best response to the signals
emitted by their peers. Our simulations replicate previous results where
the handicap principle mediates the evolution of signals. In our
simulations learning on the receiver side can mediate the evolution of
signals of quality on the sender side. When the cost of the signal is
proportional to the quality of the sender, all individuals in
populations are honest signalers. In contrast to traditional models
which predict the absence of signals when the cost is not proportional
to the quality of the signaler, our model revealed that learning
facilitates the evolution of a polymorphism in which populations
comprise both honest and dishonest signalers. We argue that learning may
have a role in the evolution and dynamics of a wide range of
communication systems and more generally in behavioral responses.
\end{abstract}
\subsection*{Lay summary}
Biological signals allow animals to respond appropriately when they
compete with each other. For example, a signal of fighting ability would
allow individuals to decide whether to fight another individual. But,
what prevents weaklings from faking to be strong? Using a theoretical
approach, we assess whether learning has a role in maintaining honest
signals. We find that when individuals learn how best to respond to a
signal, populations evolve to be composed of a mix of honest and
dishonest signals.
